\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2022-09-19 Latex/paper.tex
Fixing some words and punctuation

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function a $h$ a non-negative process. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01}), that in this case (\ref{ois_forward_rate_curve})
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta_t dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Then with a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

\subsection{FRAs Vs futures}
The cashflows in FRAs and futures are computing under different measures and consequently we have to adjust the futures price quote to transform them to FRAs price quote. As usual, we will define the forward rate  in time $t_0$ between $t_1$ and $t_2$ under the forward curve $E$ as
\begin{equation}\label{forward_rate}
L_{E}(t_0, t_1, t_2) = \frac{\left(\frac{P_{E}(t_0,t_1)}{P_{E}(t_0,t_2)} - 1 \right)}{\delta_{t_1,t_2}}
\end{equation} 
where $P_{E}(t,T)$ is the discount factor for the curve $E$ from t to T and $\delta_{t_1,t_2}$ is the year fraction between $t_1$ and $t_2$. We must observe that $L_{E}(t, t_1, t_2)$ is a martingale under the forward measure $\mathbb{Q}^{t_2}$. Let us define the future rate   
\begin{equation}\label{future}
\hat{L}_{E}(t,t_0, t_1, t_2) = \mathbb{E}_t^{\mathbb{Q}}\left(L_{E}(t_0, t_1, t_2) \right), 
\end{equation}
where the measure $\mathbb{Q}$ is the measure associated to the numeraire $B_t=\exp\left(\int_{0}^{t} r_{ois, s} ds \right)$ with $ r_{ois, t}$ the risk free short rate. From (\ref{forward_rate}) and (\ref{future}), we will define the convexity adjusment as
\begin{equation*}
CA(t, t_0, t_1, t_2) = \hat{L}_{E}(t,t_0, t_1, t_2) - \mathbb{E}_t^{\mathbb{Q}^{t_2}}\left(L_{E}(t_0, t_1, t_2) \right)
\end{equation*}
With view to get a general representation formula of the convexity adjustment for the futures, we will use (\ref{clark-okone}) to 
$L_{E}(t_0,t_1,t_2)$ i.e

\begin{equation}
L_{E}(t_0,t_1,t_2) = \hat{L}_{E}(t,t_0, t_1, t_2) + \int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s
\end{equation} 
therefore, if we take $\mathbb{E}^{\mathbb{Q}^{t2}}(\cdot)$ in the last equation, then we get
\begin{equation*}
CA(t, t_0, t_1, t_2) = - \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}^{\mathbb{Q}}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s \right).
\end{equation*}
From (\ref{ois_forward_rate_curve}) and since $f_{ois}(t,T)$ is a $\mathbb{Q}^{T}$ martingale, we have that
\begin{equation*}
dW^{\mathbb{Q}^{t_2}} = dW^{\mathbb{Q}} + \nu(t,T) dt. 
\end{equation*}
Therefore, if we apply Girsanov's theorem to switch to measure $\mathbb{Q}^{t_2}$, we get that
\begin{equation}\label{ca_general_future}
CA(t, t_0, t_1, t_2) = \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}^{\mathbb{Q}}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) \nu(s,t_2) ds \right)
\end{equation}
where $\nu(t,T)$ has been defined in (\ref{ois_forward_rate_curve}). Now, from the definition of $L_{E}(t,T)$ we have that
\begin{equation*}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)} D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right). 
\end{equation*}
If we use the representation formula (\ref{bond_ois}) we get that
\begin{equation*}
D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right) = \frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}
\end{equation*}
therefore
\begin{equation}\label{malliavin_derive_L}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}. 
\end{equation}
From (\ref{short_rate_cheyette}) y we define $\beta(s,t, x, y) = \exp\left(-\int_{s}^{t} k_{u} \right) \eta(s, x, y)$ we have that
\begin{equation}\label{malliavin_derive_x}
D_s x_{t_0} = \beta(s,t_0, x_s, y_s)M(s,t_0)
\end{equation}
where $M(s,t_0)=\exp\left(-\frac{\int_{s}^{t_0} \left(\partial_x \beta(u,t_0,x_s,y_s)\right)^{2} du}{2} + \int_{s}^{t_0} \partial_x \beta(u,t_0,x_s,y_s) dW^{\mathbb{Q}}_u \right)$. 
Then, we have that
\begin{align*}
D_s L_{E}(t_0,t_1,t_2) &= \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} \beta(s,t_0, x_s, y_s)M(s,t_0) \nonumber \\
\approx& \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)} \right)\beta(s,t_0, x_s, y_s)M(s,t_0).
\end{align*}
It is a common practice, freeze the state variables of the model to achieve some approximations. We can approximate $x_s$ and $y_s$ the next way
\begin{align*}
x_s &\approx x_0 \nonumber \\
y_s &\approx y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du
\end{align*}
Therefore, 
\begin{equation}\label{approximation_clarkocone}
\mathbb{E}_s\left( D_s L_{E}(t_0,t_1,t_2) \right) = \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)}\right)\beta(s,t_0, x_0, \hat{y}_s)
\end{equation}
where $\hat{y}_s = y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du$.
Now, from (\ref{ca_general_future}) and (\ref{approximation_clarkocone}) we have the next approximation for the convexity adjustmet for future

>>>>

In order to have a Markovian representation of the HJM, we will assume that the volatility is separable, i.e.
\begin{equation}
\sigma(t,T)= h(t)g(T).
\end{equation}
with $g$ a positive time-dependent function and $h$ a non-negative process. This version of the HJM is also known as the Cheyette model, \cite{Cheyette}. It is easy to show (see \cite{Andreasen01}), that in this case (\ref{ois_forward_rate_curve})
\begin{equation}
r_{ois}(t)=f_{ois}(t,t)= f_{ois}(0,t) + x_t,
\end{equation}  
where
\begin{align}\label{short_rate_cheyette}
dx_t &= (-k_t x_t + y_t)dt + \eta_t dW_t^{\mathbb{Q}} \nonumber \\
dy_t &= (\eta^{2}_t - 2 k_t y_t) dt ,
\end{align} 
with 
\begin{align*}
\eta_t &= g(t)h(t,x_t,y_t)  \nonumber \\
k_t &= - \frac{\partial_t g(t)}{g(t)}.
\end{align*}
Then with a little bit of algebra and we use Itô formula we can show the next representation formula for $P_{ois}(t,T)$
\begin{equation}\label{bond_ois}
P_{ois}(t,T) = \frac{P_{ois}(0,T)}{P_{ois}(0,t)} \exp\left(-G(t,T)x_t - \frac{1}{2} G^{2}(t,T)y_t \right)
\end{equation}
where $G(t,T) = \int_{t}^{T} \exp\left(-\int_{t}^{u} k_s ds \right) du$. We must observe, that under the representation (\ref{estimation_forward_rate_curve}) we have that
\begin{equation}\label{bond_forward}
P_{E}(t,T)=H(t,T)P_{ois}(t,T)
\end{equation}
where $H(t,T)=\exp\left(-\int_{t}^{T}s(t,u) du \right)$.

@@

\subsection{FRAs Vs futures}
The cashflows in FRAs and futures are computing under different measures and consequently we have to adjust the futures price quote to transform them to FRAs price quote. As usual, we will define the forward rate  in time $t_0$ between $t_1$ and $t_2$ under the forward curve $E$ as
\begin{equation}\label{forward_rate}
L_{E}(t_0, t_1, t_2) = \frac{\left(\frac{P_{E}(t_0,t_1)}{P_{E}(t_0,t_2)} - 1 \right)}{\delta_{t_1,t_2}}
\end{equation} 
where $P_{E}(t,T)$ is the discount factor for the curve $E$ from t to T and $\delta_{t_1,t_2}$ is the year fraction between $t_1$ and $t_2$. We must observe that $L_{E}(t, t_1, t_2)$ is a martingale under the forward measure $\mathbb{Q}^{t_2}$. Let us define the future rate   
\begin{equation}\label{future}
\hat{L}_{E}(t,t_0, t_1, t_2) = \mathbb{E}_t^{\mathbb{Q}}\left(L_{E}(t_0, t_1, t_2) \right), 
\end{equation}
where the measure $\mathbb{Q}$ is the measure associated to the numeraire $B_t=\exp\left(\int_{0}^{t} r_{ois, s} ds \right)$ with $ r_{ois, t}$ the risk free short rate. From (\ref{forward_rate}) and (\ref{future}), we will define the convexity adjustment as
\begin{equation*}
CA(t, t_0, t_1, t_2) = \hat{L}_{E}(t,t_0, t_1, t_2) - \mathbb{E}_t^{\mathbb{Q}^{t_2}}\left(L_{E}(t_0, t_1, t_2) \right)
\end{equation*}
With view to get a general representation formula of the convexity adjustment for the futures, we will use (\ref{clark-okone}) to 
$L_{E}(t_0,t_1,t_2)$ i.e

\begin{equation}
L_{E}(t_0,t_1,t_2) = \hat{L}_{E}(t,t_0, t_1, t_2) + \int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s.
\end{equation} 
Therefore, if we take $\mathbb{E}^{\mathbb{Q}^{t2}}(\cdot)$ in the last equation, then we get
\begin{equation*}
CA(t, t_0, t_1, t_2) = - \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}^{\mathbb{Q}}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s \right).
\end{equation*}
From (\ref{ois_forward_rate_curve}) and since $f_{ois}(t,T)$ is a $\mathbb{Q}^{T}$ martingale, we have that
\begin{equation*}
dW^{\mathbb{Q}^{t_2}} = dW^{\mathbb{Q}} + \nu(t,T) dt. 
\end{equation*}
Therefore, if we apply Girsanov's theorem to switch to measure $\mathbb{Q}^{t_2}$, we get that
\begin{equation}\label{ca_general_future}
CA(t, t_0, t_1, t_2) = \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}^{\mathbb{Q}}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) \nu(s,t_2) ds \right)
\end{equation}
where $\nu(t,T)$ has been defined in (\ref{ois_forward_rate_curve}). Now, from the definition of $L_{E}(t,T)$ we have that
\begin{equation*}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)} D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right). 
\end{equation*}
If we use the representation formula (\ref{bond_ois}) we get that
\begin{equation*}
D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right) = \frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}
\end{equation*}
therefore
\begin{equation}\label{malliavin_derive_L}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}. 
\end{equation}
From (\ref{short_rate_cheyette}) y we define $\beta(s,t, x, y) = \exp\left(-\int_{s}^{t} k_{u} \right) \eta(s, x, y)$ we have that
\begin{equation}\label{malliavin_derive_x}
D_s x_{t_0} = \beta(s,t_0, x_s, y_s)M(s,t_0)
\end{equation}
where $M(s,t_0)=\exp\left(-\frac{\int_{s}^{t_0} \left(\partial_x \beta(u,t_0,x_s,y_s)\right)^{2} du}{2} + \int_{s}^{t_0} \partial_x \beta(u,t_0,x_s,y_s) dW^{\mathbb{Q}}_u \right)$. 
Then, we have that
\begin{align*}
D_s L_{E}(t_0,t_1,t_2) &= \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} \beta(s,t_0, x_s, y_s)M(s,t_0) \nonumber \\
\approx& \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)} \right)\beta(s,t_0, x_s, y_s)M(s,t_0).
\end{align*}
It is a common practice, freeze the state variables of the model to achieve some approximations. We can approximate $x_s$ and $y_s$ the next way
\begin{align*}
x_s &\approx x_0 \nonumber \\
y_s &\approx y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du
\end{align*}
Therefore, 
\begin{equation}\label{approximation_clarkocone}
\mathbb{E}_s\left( D_s L_{E}(t_0,t_1,t_2) \right) = \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) \frac{P_{ois}(0,t_2)}{P_{ois}(0,t_0)} - G(t_0,t_1) \frac{P_{ois}(0,t_1)}{P_{ois}(0,t_0)}\right)\beta(s,t_0, x_0, \hat{y}_s)
\end{equation}
where $\hat{y}_s = y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du$.
Now, from (\ref{ca_general_future}) and (\ref{approximation_clarkocone}) we have the next approximation for the convexity adjustmet for future
